---
title: "Week2"
author: "Meng Ye"
date: "2022-08-31"
output: html_document
---

```{r load packages, message=FALSE, warning=FALSE}
#Bayes rules Chapter 2
library(bayesrules)
library(tidyverse)
library(janitor)
#Bayes rules

```

# Bayes Rules Chapter 2

## Follow along Bayes Rules Chapter 2

### Intro

```{r}
# Import article data
data(fake_news)
```

ps: data takes some to load...


```{r}
fake_news %>% 
  tabyl(type) %>% 
  adorn_totals("row")
```

ps: get to know a new way to derive the "pretend" row of summed up data, but maybe compared with our earlier `add_row()` approach, the types of calculation for the summed up row might be limited to only `sum()`.

```{r}
# Tabulate exclamation usage and article type
fake_news %>% 
  tabyl(title_has_excl, type) %>% 
  adorn_totals("row")
```



### 2.1 Building a Beyesian model for the events 


## Questions for the chapter

## Practices for Beyes Rules Chapter 2


# Bayes Rules Chapter 3

## Follow along Bayes Rules Chapter 3

## Questions for the chapter

## Practices for Beyes Rules Chapter 3






# Stats Rethinking Chapter 2

## Follow along Bayes Rules Chapter 2

## Questions for the chapter

## Practices for Beyes Rules Chapter 2






# Stats Rethinking Chapter 3

## Follow along Bayes Rules Chapter 3

## Questions for the chapter

## Practices for Beyes Rules Chapter 3













