---
title: "Ways to run Stan code"
date: "2022-09-22"
output: html_document
---


```{r}
library(rstan)
```

## Option 1: Stan code as string
Store the Stan code as a character object and then reference it with the `model_code` argument in `stan()`, which will (1) compile the code with C++ and (2) run the MCMC sampling:

```{r results="hide"}
bb_model <- "
  data {
    int<lower = 0, upper = 10> Y;
  }
  parameters {
    real<lower = 0, upper = 1> pi;
  }
  model {
    Y ~ binomial(10, pi);
    pi ~ beta(2, 2);
  }
"
bb_sim <- stan(model_code = bb_model, data = list(Y = 9), 
               chains = 4, iter = 5000*2, seed = 84735)
bb_sim
```


## Option 2: Stan code as file
Store the Stan code in a separate file on your computer (here "stan_example.stan") and then reference it with the `file` argument in `stan()`, which will (1) compile the code with C++ and (2) run the MCMC sampling:

```{r results="hide"}
bb_sim <- stan(file = "stan_example.stan", data = list(Y = 9), 
               chains = 4, iter = 5000*2, seed = 84735)
bb_sim
```

## Option 3: Stan code as chunk
Put the Stan code in a chunk and specify the the `output.var` option to store the model as an object in R. Running the Stan chunk will compile it with C++. Then use `sampling()` to run the MCMC samples (note that you don't use `stan()` here, since that both compiles and samples; running the Stan chunk compiles the model, so we only need to sample)

```{stan output.var="model_in_chunk"}
data {
  int<lower = 0, upper = 10> Y;
}
parameters {
  real<lower = 0, upper = 1> pi;
}
model {
  Y ~ binomial(10, pi);
  pi ~ beta(2, 2);
}
```
```{r}
bb_sim <- sampling(object = model_in_chunk, data = list(Y = 9), 
                   chains = 4, iter = 5000*2, seed = 84735)
bb_sim
```






