---
title: "week4_meeting"
output: html_document
date: "2022-09-22"
---



```{r packages-data, warning=FALSE, message=FALSE}
library(tidyverse)
library(brms)
library(tidybayes)
library(palmerpenguins)

penguins <- penguins %>% 
  drop_na(sex) %>% 
  mutate(flipper_length_mm_z = scale(flipper_length_mm))

flipper_scale <- attributes(penguins$flipper_length_mm_z) %>% 
  set_names(janitor::make_clean_names(names(.)))

set.seed(1234)

theme_set(theme_bw())
```



## 1: Question/goal/estimand

What is the relationship between penguin weight and penguin flipper length?

```{r}
ggplot(penguins, aes(x = flipper_length_mm, y = body_mass_g)) +
  geom_point()
```


## 2: Scientific model
Flipper length influences weight—as flippers get longer, penguins get bigger, so we have a simple DAG like this:

$$
\text{Flipper length} \rightarrow \text{Weight}
$$

We can model this by saying that penguin weight has some sort of distribution with a mean and variance. We'll pretend it's normal (it probably is too). The mean ($\mu$) of that distribution is determined in part by flipper length, and there's a linear relationship between average weight and flipper length, like this:

$$
\begin{aligned}
\text{Weight}_i &\sim \mathcal{N}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta\ \text{Flipper}_i
\end{aligned}
$$

Or with more general notation:

$$
\begin{aligned}
y_i &\sim \mathcal{N}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta x_i
\end{aligned}
$$

## 3: Statistical model

### Simulate and check the priors

We need to figure out $\alpha$, $\beta$, and $\sigma$, but we don't know them

$$
\begin{aligned}
y_i &\sim \mathcal{N}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta x_i \\
\\
\alpha &\sim \mathcal{N}(4500, 1000) \\
\beta &\sim \mathcal{N}(300, 50) \\
\sigma &\sim \operatorname{Exponential}(1)
\end{aligned}
$$

```{r}
get_prior(
  bf(body_mass_g ~ 1 + flipper_length_mm_z),
  data = penguins,
  family = gaussian()
)

priors <- c(prior(normal(4500, 750), class = Intercept),
            prior(normal(30, 50), class = b),
            prior(exponential(1), class = sigma))
```


```{r}
priors %>% 
  parse_dist() %>% 
  ggplot(aes(y = 0, dist = .dist, args = .args, fill = prior)) +
  stat_slab() +
  facet_wrap(vars(prior), scales = "free")
```


```{r}
ggplot() +
  stat_function(fun = ~dnorm(., 4500, 1000), geom = "area") +
  xlim(c(1000, 8000))

```


```{r results="hide"}
flipper_prior_only <- brm(
  bf(body_mass_g ~ 1 + flipper_length_mm_z),
  data = penguins,
  family = gaussian(),
  prior = priors,
  sample_prior = "only",
  backend = "cmdstanr"
)
```



```{r}
draws_prior <- tibble(flipper_length_mm_z = seq((170 - flipper_scale$scaled_center) / flipper_scale$scaled_scale, 
                                                (240 - flipper_scale$scaled_center) / flipper_scale$scaled_scale, 
                                                length.out = 100)) %>% 
  add_epred_draws(flipper_prior_only, ndraws = 100) %>% 
  mutate(flipper_unscaled = (flipper_length_mm_z * flipper_scale$scaled_scale) + flipper_scale$scaled_center)
draws_prior %>% 
  ggplot(aes(x = flipper_unscaled, y = .epred)) +
  geom_line(aes(group = .draw), alpha = 0.2) +
  coord_cartesian(xlim = c(170, 240), ylim = c(2500, 6000))
```


### Fit the actual model
```{r results="hide"}
flipper_model <- brm(
  bf(body_mass_g ~ 1 + flipper_length_mm_z),
  data = penguins,
  family = gaussian(),
  prior = priors,
  backend = "cmdstanr",
  chains = 4, cores = 4, seed = 1234
)
```


```{r}
flipper_model  # Means
print(flipper_model, robust = TRUE)  # Medians
```


## 4: Validate model
```{r}
pp_check(flipper_model)
```


## 5: Analyze data
```{r}
#get_variables(flipper_model)
flipper_model %>% 
  gather_draws(b_Intercept, b_flipper_length_mm_z, sigma) %>% 
  ggplot(aes(x = .value, y = fct_rev(.variable))) +
  stat_halfeye()
```


```{r}
flipper_model %>% 
  gather_draws(b_Intercept, b_flipper_length_mm_z, sigma) %>% 
  ggplot(aes(x = .value)) +
  stat_halfeye() +
  facet_wrap(vars(.variable), scales = "free")
```


```{r}
draws_posterior_epred <- tibble(flipper_length_mm_z = seq(min(penguins$flipper_length_mm_z), 
                                                          max(penguins$flipper_length_mm_z), 
                                                          length.out = 100)) %>% 
  add_epred_draws(flipper_model, ndraws = 50) %>% 
  mutate(flipper_unscaled = (flipper_length_mm_z * flipper_scale$scaled_scale) + flipper_scale$scaled_center)
ggplot() +
  geom_point(data = penguins, aes(x = flipper_length_mm, y = body_mass_g), alpha = 0.5, size = 1) +
  geom_line(data = draws_posterior_epred,
            aes(x = flipper_unscaled, y = .epred, group = .draw), alpha = 0.2)
```



```{r}
draws_posterior_pred <- tibble(flipper_length_mm_z = seq(min(penguins$flipper_length_mm_z), 
                                                         max(penguins$flipper_length_mm_z), 
                                                         length.out = 500)) %>% 
  add_predicted_draws(flipper_model, ndraws = 100) %>%
  mutate(flipper_unscaled = (flipper_length_mm_z * flipper_scale$scaled_scale) + flipper_scale$scaled_center)
ggplot() +
  geom_point(data = penguins, aes(x = flipper_length_mm, y = body_mass_g), alpha = 0.5, size = 1) +
  stat_lineribbon(data = draws_posterior_pred,
                  aes(x = flipper_unscaled, y = .prediction), .width = 0.89, alpha = 0.2)
```

















