---
title: "Week2 Rethinking"
author: "Meng Ye"
date: "2022-09-02"
output: html_document
---

```{r load packages, message=FALSE, warning=FALSE}
library(bayesrules)
library(tidyverse)
library(rethinking)
library(patchwork)
```


# Stats Rethinking Chapter 2

## Notes and selected follow along

### 2.1 The Garden of Forking Data

No superstition. Bayesian inference is just counting and comparing possibilities. 

Possibilities -> *conjectures*

Principle of indifference, comparable to non-Bayesian approaches. The structure of the model and the scientific context always provide information that allows us to do better than ignorance.



### 2.2 Building a Model

A Bayesian model begins with one set of plausibilities assigned to each of these possibilities. Then it updates them in light of the data, to produce the posterior plausibilities. 

It is possible to mathematically divide out the observation, to infer the previous plausibility curve. So the data could be presented to your model in any order, or all at once even. 

Non-Bayesian care for sample size -> In contrast, Bayesian estimates are valid for any sample size. But the price for this power is dependency upon the initial plausibilities, the prior. 


### 2.3 Components of the Model

information entropy/maximum entropy: disorder? death of the universe? Actually it eans that the distribution contains no additional information other than: two events, with p and 1-p


- observed variables: the likelihood function part

- unobserved variables: the parameters. "Since p is not observed, we usually call it a parameter". 

Bayes' theorem:

$$Posterior = \frac{Probility\;of\; the\; data \times Prior}{Average\; Probability\; of\; the\; data}$$

Average probability: average over the prior, its job is just to standardize the posterior

Bayes rules: Law of total probability

Also the expectation or *marginal likelihood*


### 2.4 Making the Model Go

data (likelihood) + prior (of parameters) -> posterior distribution

The distribution contains the relative plausibility of different parameter values, conditional on the data and the model. 

Three conditioning engines:

- (1) Grid approximation
- (2) Quadratic approximation
- (3) Markov chain Monte Carlo (MCMC)

### example: the grid approximation method

```{r}
set.seed(5431)
p_grid <- seq(0, 1, length.out = 20)

```

```{r}
# Probability of each value of p
# Super vague uniform prior: just 1 at each possible p
prob_p_uniform <- rep(1, 20)

# Probability of each proportion, given 6/9 water draws
prob_data <- dbinom(6, size = 9, prob = p_grid)

# Unnormalized posterior
posterior_raw <- prob_data * prob_p_uniform

# Normalized posterior that sums to 1
posterior_normalized <- posterior_raw / sum(posterior_raw)

plot(p_grid, posterior_normalized, type = "b")
```
tidyverse practice: replication of Figure 2.6 

```{r}
sequence_length <- 1e3

data_multiple <-
  tibble(probability = seq(from = 0, to = 1, length.out = sequence_length)) %>% 
  expand(probability, row = c("flat", "stepped", "Laplace"))%>% 
  arrange(row, probability) %>% 
  mutate(prior = ifelse(row == "flat", 1,
                        ifelse(row == "stepped", rep(0:1, each = sequence_length / 2),
                               exp(-abs(probability - 0.5) / .25) / ( 2 * 0.25))),
         likelihood = dbinom(x = 6, size = 9, prob = probability)) %>% 
  group_by(row) %>% 
  mutate(posterior = prior * likelihood / sum(prior * likelihood)) %>% 
  pivot_longer(prior:posterior)  %>% 
  ungroup() %>% 
  mutate(name = factor(name, levels = c("prior", "likelihood", "posterior")),
         row  = factor(row, levels = c("flat", "stepped", "Laplace")))
```


```{r}
p1 <-
  data_multiple %>%
  filter(row == "flat") %>% 
  ggplot(aes(x = probability, y = value)) +
  geom_line() +
  scale_x_continuous(NULL, breaks = NULL) +
  scale_y_continuous(NULL, breaks = NULL) +
  theme(panel.grid = element_blank()) +
  facet_wrap(~ name, scales = "free_y")

p2 <-
  data_multiple %>%
  filter(row == "stepped") %>% 
  ggplot(aes(x = probability, y = value)) +
  geom_line() +
  scale_x_continuous(NULL, breaks = NULL) +
  scale_y_continuous(NULL, breaks = NULL) +
  theme(panel.grid = element_blank(),
        strip.background = element_blank(),
        strip.text = element_blank()) +
  facet_wrap(~ name, scales = "free_y")

p3 <-
  data_multiple %>%
  filter(row == "Laplace") %>% 
  ggplot(aes(x = probability, y = value)) +
  geom_line() +
  scale_x_continuous(NULL, breaks = c(0, .5, 1)) +
  scale_y_continuous(NULL, breaks = NULL) +
  theme(panel.grid = element_blank(),
        strip.background = element_blank(),
        strip.text = element_blank()) +
  facet_wrap(~ name, scales = "free_y")

# combine
p1 / p2 / p3
```

Read the explanation for quadratic approximation and MCMC for intuition for now. Save writing codes to replicate the process later. 
ps: rationable behind R code 2.7 indeed was explained in Bayes rules.

## Questions for the chapter

1. Bayes rules call the updating information the likelihood function, the Rethinking calls it  ways the outcome observed can produce. Apart from the intuition, can I understand this term as the weighting adjustment of the prior to get the posterior?

2. In Bayesian analysis, the coef for the main IV is the "parameter", how about the coef for control variables?

3. The normalizing constant: both explanation as average probability and the law of total probability is correct (weighted average)？


$$Pr(W,L) = E (Pr(W,L|p)) = \int Pr(W,L|p) Pr(p)dp$$

4. About the parser and denser grid exemplified on P41, denser grid does better approximation (the marginal return drops), so why does the author say sample size does not matter for Bayesian? 

## Practices 

### 2H1

Understanding the question as:

p(twins) prior distribution is 0.5/0.5 (equally common), and p's values are 0.1 and 0.3.

If we understand her next birth as an independent event, what 2H1 asks is the marginal probability that is 

$$marginal\; probability = 0.1*0.5+0.2*0.5=0.15$$ 
### 2H2



```{r}
#still borrowing Bayes rules terming of pi
panda <- tibble(pi_value = c(0.1, 0.2),
                pi_prior = c(0.5, 0.5)) %>% 
  mutate(likelihood = choose(1,1) * pi_value^1 * (1 - pi_value)^0) %>%
  mutate(unnormalized = pi_prior * likelihood) %>% 
  mutate(normalizing_constant = sum(unnormalized)) %>% 
  mutate(pi_posterior = unnormalized/normalizing_constant)

panda$pi_posterior
```

The probability that the panda mother is from Species A is 0.3333.


### 2H3
Updating the likelihood function 

```{r}
#still borrowing Bayes rules terming of pi
panda2 <- tibble(pi_value = c(0.1, 0.2),
                pi_prior = c(0.5, 0.5)) %>% 
  mutate(likelihood = choose(2,1) * pi_value^1 * (1 - pi_value)^1) %>%
  mutate(unnormalized = pi_prior * likelihood) %>% 
  mutate(normalizing_constant = sum(unnormalized)) %>% 
  mutate(pi_posterior = unnormalized/normalizing_constant)

panda2$pi_posterior
```

The probability that the panda mother is from Species A is 0.36.


### 2H4
First without birth data

```{r}
#still borrowing Bayes rules terming of pi
panda_test <- tibble(pi_value = c(0.8, 0.65),
                pi_prior = c(0.5, 0.5)) %>% 
  mutate(likelihood = choose(1,1) * pi_value^1 * (1 - pi_value)^0) %>%
  mutate(unnormalized = pi_prior * likelihood) %>% 
  mutate(normalizing_constant = sum(unnormalized)) %>% 
  mutate(pi_posterior = unnormalized/normalizing_constant)

panda_test$pi_posterior
```

Then with birth data Twin + Single 

```{r}
#still borrowing Bayes rules terming of pi
panda_test2 <- tibble(pi_value = c(0.8, 0.65),
                pi_prior = c(0.36, 0.64)) %>% 
  mutate(likelihood = choose(1,1) * pi_value^1 * (1 - pi_value)^0) %>%
  mutate(unnormalized = pi_prior * likelihood) %>% 
  mutate(normalizing_constant = sum(unnormalized)) %>% 
  mutate(pi_posterior = unnormalized/normalizing_constant)

panda_test2$pi_posterior
```

# Stats Rethinking Chapter 3

## Notes and selected follow along

### Intro

- probability format: harder to understand + present frequencies of events rather than theoretical parameters, all major statistical philosophies would agree to use Bayes' theorem in this case. In other words, Bayesian statistics $\neq$ Bayes Theorem

- using counts rather than probabilities, *frequency format* or *natural frequencies*

- imagine the probability distribution as drawing samples from it. The sampled events in this case are **parameter values**. This chapter: working with samples from the **posterior distribution**.

- The Bayesian formalism treats parameter distributions as *relative plausibility*, not as any physical random process.

- Why sampling? Transform integral (total probability in some interval) to counting values in the interval. Working with samples is enough to answer a bunch of question, without relying upon a captive mathematician. (Meaning approximating counious prob distribution with discrete ones?)

### 3.1 Sampling from a Grid-Approximate posterior

```{r}
p_grid_ch3 <- seq(0, 1, length.out = 1000)

prob_p <- rep(1, 1000)
prob_data <- dbinom(6, size = 9, prob = p_grid_ch3)
posterior_raw_ch3 <- prob_data * prob_p
posterior_normalized_ch3 <- posterior_raw_ch3 / sum(posterior_raw_ch3)

samples <- sample(p_grid_ch3, prob = posterior_normalized_ch3, size = 1e4, replace = TRUE)
plot(samples)
```
```{r}
# R code 3.5
dens(samples)
```

All you've done is crudely replicate the posterior density you **had already computed**. This isn't of much value.


### 3.2 Sampling to Summarize

How to summarize posterior distribution depends on your purpose. 

- intervals of defined boundaries
- intervals of defined probability mass 
  percentile intervals -> HPDI 
  compatibility intervals: communicate the shape of a distribution
- point estimates 


```{r}
# R code 3.11
set.seed(5321)
prob_data_skewed <- dbinom(3, size = 3, prob = p_grid_ch3)
posterior_raw_skewed <- prob_data_skewed * prob_p
posterior_normalized_skewed <- posterior_raw_skewed / sum(posterior_raw_skewed)

samples_skewed <- sample(p_grid_ch3, prob = posterior_normalized_skewed, size = 1e4, replace = TRUE)
```

```{r}
# similar result to R code 3.12
PI(samples, prob = 0.5)
```

```{r}
HPDI(samples_skewed, prob = 0.5)
```

The HPDI is narrower, according to the author, it is the narrowest by defination. 

Remember, the entire posterior distribution is the Bayesian "estimate". It summarizes the relative plausibilities of each possible value of the parameter. 

### 3.3 Sampling to Simulate Prediction

## Questions for the chapter

1. What are the differences of `slice_sample()`, `sample_n()`, and `sample()`? In other words, how do I know which command to choose under various circumstances?

2. In real Bayesian stats, we mostly don't know the posterior distribution, just the approximation right?


## Practices 















