---
title: "Week4_rethinking"
author: "Meng Ye"
date: "2022-09-19"
output: html_document
---

```{r load packages, message=FALSE, warning=FALSE}
library(tidyverse)
library(rethinking)
```


# Chapter 5 spurious waffle 

That is pivoting from simple linear regeression to multiple regression 

## Intro

Reasons for multiple regressions 

1) control for confounders 

2) complex causal mechanisms - multiple causes works simultaneously 

3) Interaction

Also, causal inference, DAG, over-controlling bias, e.g. collider bias 

## 5.1 Spurious Association 

```{r}
library(dagitty) 
dag5.1 <- dagitty( "dag{ A -> D; A -> M; M -> D}")  
coordinates(dag5.1) <- list( x=c(A=0,D=1,M=2), y=c(A=0,D=1,M=0))  
drawdag(dag5.1) 
```



# Draft discussion agenda for Thursday 

1. Basic concepts/elements in Bayesian linear regression, $\mu_i$, comparing to OLS. Beta is the same variable? The output from `brm()` is also the mean or the MAP? 

https://andrewheiss.slack.com/archives/D02CATRD4JZ/p1662742163766179


2. Understanding the underlying mechanisms for Bayesian linear regression posterior calculation (based on content in previous weeks), the simulation example

https://andrewheiss.slack.com/archives/D02CATRD4JZ/p1663045563153539

3. Compare the purposes and functions of different steps in Bayesian model fitting
   a. simulate the prior predictive distribution p95 
   b. fitting model, either `quap()` or `brm()`
   c. presenting model outputs: 
    `rethinking::precis()` - tables of "marginal distribution"?     `rethinking::vcov()` var-cov matrics for the parameters, what for? 
    plotting regression intervals and contours, the average regression line? How about the MAP of the joint probability distribution of $\alpha$ and $\beta$
    what outputs does `brm()` gives? read one ouput example togeter
    
  https://andrewheiss.slack.com/archives/D02CATRD4JZ/p1663601103979599
  

4. `brm()` syntax details: the options of "cores", "chains" ... 

